\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\setstretch{1.15}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\begin{document}

\begin{center}
{\Large\textbf{Response to Technical Check}}\\[0.6em]
{\normalsize Journal of Family and Economic Issues}\\[1.2em]
\rule{0.82\textwidth}{0.4pt}
\end{center}

\vspace{0.6em}

\textbf{Manuscript Title:} Financial Preparedness, Reproductive Education, and Completed Fertility: Evidence From the NLSY79\\
\textbf{Manuscript ID:} 62043d98-011b-493b-a278-acdf3363ea72 v1.0

\vspace{0.8em}

Dear Ramesh Anandhan and the Editorial Support Team,

Thank you for the prompt technical check of our manuscript. We are grateful for the fast turnaround time and for the care taken in reviewing our first paper for the \textit{Journal of Family and Economic Issues}.

We have addressed the requested technical-check item carefully. The author declaration material has been removed from the blinded manuscript file. Funding, competing-interest, author-contribution, ethics, consent, and related declaration information will be provided through the Snapp submission fields or other administrative fields requested by the journal, rather than included in files shared with reviewers.

The reviewer-facing source package now contains only the anonymized manuscript source file and the references file. It does not include a PDF, title page, funding statement file, declarations file, cover letter, or other author-identifying submission material.

No scholarly manuscript-body changes were made for this technical-check return.

Thank you again for your time, guidance, and assistance as the manuscript moves through the review process.

\vspace{1.2em}

Sincerely,\\[0.6em]
Submitting Author

\end{document}
