\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage[hidelinks]{hyperref}
\doublespacing

\begin{document}

% ************************************************************
% TITLE PAGE AND DECLARATIONS
% ************************************************************

\begin{center}
{\Large Financial Preparedness, Birth-Control Education, and Family Formation in the NLSY79}\par
\vspace{1em}
Paul A. Skeffington\par
Dartmouth College\par
\end{center}

\vspace{1em}

\section*{Declarations}

\subsection*{Funding}

No funding was received to assist with the preparation of this manuscript. No funds, grants, or other support were received for conducting this study.

\subsection*{Competing Interests}

The author declares no competing interests.

\subsection*{Ethics Approval}

This study uses secondary, de-identified public-use data from the National Longitudinal Survey of Youth 1979 (NLSY79). No new human-subjects data were collected, and the study did not involve direct interaction with participants. Because the analysis relies exclusively on de-identified secondary data, additional institutional ethics approval was not required for this manuscript.

\subsection*{Consent to Participate}

Not applicable. This study uses secondary, de-identified public-use survey data. No new participant recruitment, contact, intervention, or consent procedure was conducted for this analysis.

\subsection*{Consent for Publication}

Not applicable. The manuscript does not contain personally identifiable respondent information, individual case narratives, images, or other materials requiring separate consent for publication.

\subsection*{Data Availability}

The data used in this study are from the National Longitudinal Survey of Youth 1979, administered by the U.S. Bureau of Labor Statistics. Public-use NLSY79 data and documentation are available through the National Longitudinal Surveys data portal and NLS Investigator. The analytic dataset was constructed from selected public-use variables measuring marital status, fertility, emergency savings, financial literacy, and birth-control education. No restricted-use geographic identifiers, confidential respondent identifiers, or protected respondent-level information were used. Replication is subject to the availability of the public-use NLSY79 files, the selected variable extraction, and the data-cleaning procedures described in the manuscript.

\subsection*{Code Availability}

Replication code and derived analytic files may be made available upon reasonable request, subject to the structure of the public-use NLSY79 extraction, the selected variable set, and any applicable data-use restrictions from the National Longitudinal Surveys program.

\subsection*{Author Contributions}

Paul A. Skeffington designed the study, developed the research question, reviewed the literature, constructed the analytic framework, prepared the analytic dataset, conducted the statistical analysis, interpreted the results, and wrote the manuscript.

\subsection*{Acknowledgements}

The author acknowledges the U.S. Bureau of Labor Statistics and the National Longitudinal Surveys program for maintaining and providing access to the NLSY79 public-use data. The author also acknowledges Dartmouth College for academic affiliation and research development support.

% ************************************************************
% END TITLE PAGE AND DECLARATIONS
% ************************************************************

\end{document}
