% ============================================================
% TITLE PAGE / UNBLINDED SUBMISSION FILE
% Journal of Family and Economic Issues
% ============================================================

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage[hidelinks]{hyperref}

\doublespacing

\begin{document}

\begin{center}
{\Large \textbf{Financial Preparedness, Reproductive Education, and Completed Fertility: Evidence From the NLSY79}}\\[1.5em]

{\large Paul A. Skeffington}\\[0.75em]

Southern New Hampshire University\\
Dartmouth College\\[0.75em]

Corresponding author: Paul A. Skeffington\\
Email: \href{mailto:paul@skeffington.us}{paul@skeffington.us}\\
ORCID: \href{https://orcid.org/0009-0009-7786-5903}{0009-0009-7786-5903}
\end{center}

\vspace{2em}

\section*{Author Information}

Paul A. Skeffington is affiliated with Southern New Hampshire University and Dartmouth College.

\section*{Corresponding Author}

Paul A. Skeffington\\
Email: \href{mailto:paul@skeffington.us}{paul@skeffington.us}\\
ORCID: \href{https://orcid.org/0009-0009-7786-5903}{0009-0009-7786-5903}

\section*{Funding}

The author received no financial support for the research, authorship, or publication of this article.

\section*{Competing Interests}

The author declares no competing interests.

\section*{Ethics Approval}

This study used de-identified public-use data from the National Longitudinal Survey of Youth 1979. No human subjects research involving identifiable private information was conducted by the author for this manuscript.

\section*{Consent to Participate}

Not applicable. The analysis used secondary de-identified public-use survey data.

\section*{Consent for Publication}

Not applicable. No individual-level identifiable information is reported.

\section*{Data Availability}

The NLSY79 public-use data are available from the U.S. Bureau of Labor Statistics and the NLS Investigator platform. The analytic code and manuscript materials are available in the project repository, subject to public-use data redistribution rules and repository documentation.

\section*{Author Contributions}

Paul A. Skeffington was responsible for conceptualization, data construction, analysis, interpretation, and manuscript preparation.

\end{document}
